CrypTool Version 1.3 \\ Tutorial

### 2001-11-27 script-de/script-de.tex
no message

{\href{http://www.math.Princeton.EDU/~arbooker/nthprime.html}{\tt
http://www.math.Princeton.EDU/\~arbooker/nthprime.html}}

@@

{\href{http://www.math.Princeton.EDU/~arbooker/nthprime.html}{http://www.math.Princeton.EDU/\~arbooker/nthprime.html}}
\vskip +7 pt Ausgabe der n-ten Primzahl

@@

\begin{enumerate}
   \item \href{http://www.utm.edu/}{\tt http://www.utm.edu/ }
   \item \href{http://prothsearch.net/index.html}{\tt http://prothsearch.net/index.html}
   \item \href{http://www.mersenne.org/prime.htm}{\tt http://www.mersenne.org/prime.htm }
   \item \href{http://reality.sgi.com/chongo/prime/prime_press.html}{\tt http://reality.sgi.com/chongo/prime/prime\_press.html }
   \item \href{http://www.eff.org/coop-awards/prime-release1.html}{\tt http://www.eff.org/coop-awards/prime-release1.html }
%   \item \href{http://www.octocad.demon.co.uk/mersenne/prime.htm}{\tt http://www.octocad.demon.co.uk/mersenne/prime.htm }
   \item \href{http://www.informatik.tu-darmstadt.de/TI/LiDIA/}{\tt http://www.informatik.tu-darmstadt.de/TI/LiDIA/ }
   \item \href{http://www.math.Princeton.EDU/~arbooker/nthprime.html}{\tt http://www.math.Princeton.EDU/\~arbooker/nthprime.html }
   \item \href{http://www.utm.edu/research/primes/notes/by_year.html}{\tt http://www.utm.edu/research/primes/notes/by\_year.html }
   \item \href{http://www.cerias.purdue.edu/homes/ssw/cun}{\tt http://www.cerias.purdue.edu/homes/ssw/cun }
\end{enumerate}

>>>>

{\href{http://www.math.Princeton.EDU/~arbooker/nthprime.html}{\tt
http://www.math.Princeton.EDU/\~{}arbooker/nthprime.html}}

@@

{\href{http://www.math.Princeton.EDU/~arbooker/nthprime.html}{http://www.math.Princeton.EDU/\~{}arbooker/nthprime.html}}
\vskip +7 pt Ausgabe der n-ten Primzahl

@@

\begin{enumerate}
   \item \href{http://www.utm.edu/}{\tt http://www.utm.edu/ }
   \item \href{http://prothsearch.net/index.html}{\tt http://prothsearch.net/index.html}
   \item \href{http://www.mersenne.org/prime.htm}{\tt http://www.mersenne.org/prime.htm }
   \item \href{http://reality.sgi.com/chongo/prime/prime_press.html}{\tt http://reality.sgi.com/chongo/prime/prime\_press.html }
   \item \href{http://www.eff.org/coop-awards/prime-release1.html}{\tt http://www.eff.org/coop-awards/prime-release1.html }
%   \item \href{http://www.octocad.demon.co.uk/mersenne/prime.htm}{\tt http://www.octocad.demon.co.uk/mersenne/prime.htm }
   \item \href{http://www.informatik.tu-darmstadt.de/TI/LiDIA/}{\tt http://www.informatik.tu-darmstadt.de/TI/LiDIA/ }
   \item \href{http://www.math.Princeton.EDU/~arbooker/nthprime.html}{\tt http://www.math.Princeton.EDU/\~{}arbooker/nthprime.html }
   \item \href{http://www.utm.edu/research/primes/notes/by_year.html}{\tt http://www.utm.edu/research/primes/notes/by\_year.html }
   \item \href{http://www.cerias.purdue.edu/homes/ssw/cun}{\tt http://www.cerias.purdue.edu/homes/ssw/cun }
\end{enumerate}

### 2001-11-27 script-gb/script-gb.tex
no message

Every whole number has a factor. The number 1 only has one factor, itself,
whereas the number $12$ has the six factors $1, 2, 3, 4, 6, 12$. Many numbers
can only be divided by themselves and by $1$. With respect to multiplication,
these are the \glqq atoms\grqq {} in the area of numbers. Such numbers are
called prime numbers.

@@

The largest prime numbers known today have several hundred thousand digits,
which is too big for us to imagine. The number of elementary particles in the
universe is \glqq only\grqq\ estimated to be a $80$-digit number.
\hyperlink{grosord}{(See: overview under 2.6.3 of various orders of magnitude /
dimensions.)}

@@

As early as in the 19th century, it was discovered that the $29$-digit number $$
F(7) = 2^{2^7} + 1 $$ is not prime. However, it was not until 1970 that
Morrison/Billhart managed to factorise it.
\begin{eqnarray*}
F(7) & = & 34,279,974,696,877,740,253,374,607,431,768,211,457 \\
& = & 59, 649, 589, 127, 497, 217 \cdot  574, 689, 200, 685,
129,054, 721
\end{eqnarray*}
Many rapid prime number tests are based on the (small) Fermat theorem put
forward by Fermat in 1640 .
\index{Fermat!small Fermat}
\begin{theorem}[\glqq small\grqq\ Fermat]
Let $p$ be a prime number and $a$ be any whole number, then for all $a$ $$a^p
\equiv a \; {\rm mod} \; p$$
This could also be formulated as follows: \\ Let $p$ be a prime number and $a$
be any whole number that is not a multiple of $p$ (also $a \not\equiv 0 \; {\rm
mod} \; p$), then $a^{p-1} \equiv 1 \; {\rm mod} \; p$.
\end{theorem}

@@

{\href{http://www.math.Princeton.EDU/~arbooker/nthprime.html}{http://www.math.Pr
inceton.EDU/\~arbooker/nthprime.html}} \vskip +7 pt Output of the
nth prime number

@@

\begin{enumerate}
   \item \href{http://www.utm.edu/}{\tt http://www.utm.edu/ }
   \item
\href{http://prothsearch.net/index.html}{\tt
http://prothsearch.net/index.html}
   \item \href{http://www.mersenne.org/prime.htm}{\tt
http://www.mersenne.org/prime.htm }
   \item \href{http://reality.sgi.com/chongo/prime/prime_press.html}{\tt
http://reality.sgi.com/chongo/prime/prime\_press.html }
   \item \href{http://www.eff.org/coop-awards/prime-release1.html}{\tt
http://www.eff.org/coop-awards/prime-release1.html }
\item \href{http://www.eff.org/coop-awards/prime-release1.html}{\tt
%http://www.octocad.demon.co.uk/mersenne/prime.htm }
%   \item \href{http://www.informatik.tu-darmstadt.de/TI/LiDIA/}{\tt
http://www.informatik.tu-darmstadt.de/TI/LiDIA/ }
   \item \href{http://www.math.Princeton.EDU/~arbooker/nthprime.html}{\tt
http://www.math.Princeton.EDU/\~arbooker/nthprime.html }
   \item \href{http://www.utm.edu/research/primes/notes/by_year.html}{\tt
http://www.utm.edu/research/primes/notes/by\_year.html }
   \item \href{http://www.cerias.purdue.edu/homes/ssw/cun}{\tt
http://www.cerias.purdue.edu/homes/ssw/cun }
\end{enumerate}

@@

Another name for this type of attack is \index{Man-in-the-middle attack} \glqq
man-­in-­the-­middle attack\grqq. Users are promised protection against this
type of attack by public­key certification, which is intended to guarantee the
\index{Authenticity} authenticity of public keys. The most common certification
method is the X.509 standard.

@@

\paragraph{Cyclic groups}
Cyclic groups are those groups $G'$ that possess an element $g$ from which the
group operation can be used to generate all other elements in the group. This
means that for each element $a$ in $G'$ there exists a positive whole number $i$
such that if $g$ is subject to the operation $i$ times (i.e. \glqq$g \cdot
i$\grqq), $g+g+...+g = a$ (additive group) or $g^i = g\cdot g \cdots g = a$
(multiplicative group). The element $g$ is the {\em generator} of the cyclic
group --- each element in $G´$ can be generated using $g$ and the operation.

@@

There are {\em infinite} fields, i.e. the set on which the field is based
contains an infinite number of elements (e.g.: the field of real numbers). And
there are also finite fields, such as ${\mathbb Z}_p = \{0, 1, 2, 3, ..., p-1\}$
, where $p$ is a prime. ${\mathbb Z}_p$ with addition mod $p$ and multiplication
mod $p$ is a finite field.
\index{Field!Characteristic}
\paragraph{Characteristic of a field}
Let $K$ be a field and $1$ be the neutral element of $K$ with respect to the
multiplicative operation \glqq$\cdot$\grqq. For positive natural numbers $n$,
let us understand $n_1$ to be $n_1 = 1 + 1 + ... + 1$ ($n$ summands and $n_1$ is
an element in $K$). If $n_1$ is then unequal to $0$ for all $n>0$, then we call
$K$ a field with characteristic zero. Otherwise, the characteristic of $K$ is
defined to be the smallest positive natural number $p$ for which $p_1 = 0$
(note: $p$ is then a prime). Comment: The field of real numbers has the
characteristic $0$; the field ${\mathbb Z}_p$ has the characteristic $p$.

@@

An elliptic curve over ${\mathbb Z}_p$ is defined by an equation of the
following form: $$ y^2 \; ({\rm mod} \; p) = x^3 + ax + b ({\rm mod} \; p) $$
(thus: equality in the field ${\mathbb Z}_p$), where $a, b$ are in ${\mathbb
Z}_p$ and $4a^3 + 27b^2$ mod $p$ is not equal to zero. For fixed chosen numbers
$a$ and $b$ in ${\mathbb Z}_p$, this equation has the pair of solutions $$ {\bf
E} = \left\{(x,y) \left| \begin{array}{c} x {\rm ~and~} y {\rm ~are~in~}
{\mathbb Z}_p {\rm ~and~} \\ y^2  \equiv x^3 + ax + b \; ({\rm mod~} p) {\rm
~and~} \\ 4a^3 + 27b^2 \not\equiv 0 \; ({\rm mod} p)\end{array} \right.
\right\}, $$ i.e. the set ${\bf E}$ consists of all pairs $x$ and $y$ that are a
solution (in ${\mathbb Z}_p$) of the above equation. It must be noted that the
numbers $a, b$ and $p$ determine which pairs $(x,y)$ lie in the set ${\bf E}$.
This means that $a, b$ and $p$ specify this set. The elements $(x,y)$ in ${\bf
E}$ are called points on the elliptic curve. In addition, ${\bf E}$ has one more
element $O$ (the so-called point in infinity). The set ${\bf E}$ is usually
called an elliptic curve. We can now define an operation (also denoted as $+$,
although it is not the standard/usual addition of real numbers) on two elements
in ${\bf E}$ such that the operation delivers an element that also lies in ${\bf
E}$. The set ${\bf E}$ is therefore closed under the operation $+$. We can show
that ${\bf E}$ is a group. The neutral element of the group ${\bf E}$ is the
point in infinity $O$. Thus, for every two points $(x_1,y_1)$ and $(x_2,y_2)$ on
the elliptic curve ${\bf E}$, there exists a point $(x_3,y_3)$ on ${\bf E}$ such
that the operation $+$ complies with the following: $(x_1,y_1) + (x_2,y_2) =
(x_3,y_3)$. Under certain circumstances, these points may also be equal to the
point in infinity. Thus, when we speak of a point $P$ on an elliptic curve ${\bf
E}$, we mean that $P = (x,y)$ and $(x,y)$ lies in the set ${\bf E}$. Any two
points on an elliptic curve specified by $a, b$ and $p$ can therefore be added
and the result is a point that also lies on the same elliptic curve. We must
note that ${\bf E}$ can have the following meanings:
\begin{itemize}
   \item the set ${\bf E}$ of solutions pairs $(x,y)$ for an equation including the point $O$
   \item the group ${\bf E}$ (with the operation \glqq addition of $(x_1,y_1)$ and $(x_2,y_2)$\grqq)
   \item the elliptic curve ${\bf E}$ (which is actually the same as the group ${\bf E}$)
 \end{itemize}

@@

For cryptography, the important fact is that, for very large numbers, it appears
to be extremely difficult to use a given point $Q$ on an elliptic curve to
determine which two points have to be added to obtain $Q$. For large numbers $a,
b$ and $p$ ($p$, for example, has a length of more than $160$ bits), the
computer can easily add the point $P$  $m$ times after another, i.e. to
determine the point $P + P + \cdots + P = Q$ in an incredibly short space of time (in a few
fractions of a second) ($m$ summands $P$). Rather than $P + P + \cdots + P
= Q$ ($m$ summands $P$) we also write $mP = Q$. If we have a point $P$ and a
point $Q$, which both lie on the same elliptic curve, no procedure is known that
enables us --- within an acceptable space of time --- to determine the number
$m$ (assuming it actually exists) for which $mP = Q$. This is referred to as the
\glqq elliptic curve discrete logarithm problem\grqq\ (abbreviated to ECDLP).

We must note that not all elliptic curves are equally secure. This means that we
must choose the parameters $a$ and $b$ carefully when defining a curve. For
certain classes of elliptic curves, it is possible to solve the ECDLP easier
than in the general case. Cryptographically unsuitable elliptic curves are
called abnormal curves (these are curves over ${\mathbb Z}_p$ for which the set
${\bf E}$ has precisely $p$ elements) and the supersingular curves (the curves
for which we can reduce the calculation of the ECDLP to calculating the \glqq
standard\grqq\ discrete logarithm in other finite fields, i.e. simplify the
calculation). There are therefore cryptographically good and bad curves.
However, for given parameters $a$ and $b$ we can, with some difficulty,
establish whether or not the resulting elliptic curve is useful for
cryptographic purposes. The curves used in cryptography are usually provided by
experts. They ensure that the elliptic curves they classify as secure satisfy
the current security requirements. For secure curves, the parameter $p$
determines how long it takes to solve the ECLDP on this curve. The larger the
parameter $p$, the longer it takes to solve the problem. Experts recommend a bit
length of over $200$ bits for the parameter $p$. This makes it clear why
elliptic curves are so interesting for cryptography. Because the parameter $p$
also determines the time required to perform the signature/encryption procedure
when using elliptic curves in cryptography. The time taken to generate a pair of
keys also depends on $p$. Thus, small values (few bits) are desirable here (in
order to minimise the run times for the procedures); however, the required
security must still be maintained. For example, with a length of $200$ bits for
$p$, a {\em good} elliptic curve is just as secure as an \index{RSA!module} RSA
module of over $1024$ bits in length (at least according to the current state of
research). The reason for this is that the quickest algorithms for solving the
{\em elliptic curve discrete logarithm} problem have an exponential run time ---
unlike the subexponential run times that the best factorisation algorithms
currently have (number sieve, quadratic sieve or factorisation with elliptic
curves). Therefore, the parameters for cryptographic procedures based on the
problem of {\em factorising whole numbers} must be greater than the parameters
for cryptographic procedures based on the ECDL problem.

@@

The {\em elliptic curve discrete logarithm problem} (ECLDP) forms the basis for
elliptic-curve cryptography. Various signature procedures are based on this.
What they have in common is how they use the public parameters and how these
parameters are used to generate secret and public keys:
\begin{itemize}
    \item The parameters of the elliptic curve E, i.e. a prime number $p$ that
determines over which field ${\mathbb Z}_p$ the elliptic curve ${\bf E}$ is
defined, as well as the two numbers $a$ and $b$ in ${\mathbb Z}_p$.
    \item A point $G=(x,y)$ that lies on the elliptic curve ${\bf E}$.
    \item A prime number $r < p$ for which $rG=O$ (i.e. $G$ added $r$ times
gives the neutral element of the group ${\bf E}$) and $r$ is a factor of $\#{\bf
E}$ (where $\# {\bf E}$ is the number of elements in the set ${\bf E}$). The
point $G$ therefore has the order $r$ and is the generator of a cyclic subgroup
of ${\bf E}$ with the order $r$.
    \item The number $k = \#{\bf E}/r$ ($k$ is called the {\em cofactor}).
\end{itemize}
The parameters $a, b, p, G, r$ and $k$ listed above are called domain
parameters. They determine on which elliptic curve ${\bf E}$ and in which cyclic
subgroup of ${\bf E}$ a signature procedure has been \glqq used\grqq\.

>>>>

Every whole number has a factor. The number 1 only has one factor,
itself, whereas the number $12$ has the six factors $1, 2, 3, 4,
6, 12$. Many numbers can only be divided by themselves and by $1$.
With respect to multiplication, these are the ''atoms'' in the
area of numbers. Such numbers are called prime numbers.

@@

The largest prime numbers known today have several hundred
thousand digits, which is too big for us to imagine. The number of
elementary particles in the universe is ''only'' estimated to be a
$80$-digit number. \hyperlink{grosord}{(See: overview under 2.6.3
of various orders of magnitude / dimensions.)}

@@

As early as in the 19th century, it was discovered that the $29$-digit number $$
F(7) = 2^{2^7} + 1 $$ is not prime. However, it was not until 1970 that
Morrison/Billhart managed to factorise it.
\begin{eqnarray*}
F(7) & = & 34,279,974,696,877,740,253,374,607,431,768,211,457 \\
& = & 59, 649, 589, 127, 497, 217 \cdot  574, 689, 200, 685,
129,054, 721
\end{eqnarray*}
Many rapid prime number tests are based on the (small) Fermat theorem put
forward by Fermat in 1640 .
\index{Fermat!small Fermat}
\begin{theorem}[''small'' Fermat]
Let $p$ be a prime number and $a$ be any whole number, then for all $a$ $$a^p
\equiv a \; {\rm mod} \; p$$
This could also be formulated as follows: \\ Let $p$ be a prime number and $a$
be any whole number that is not a multiple of $p$ (also $a \not\equiv 0 \; {\rm
mod} \; p$), then $a^{p-1} \equiv 1 \; {\rm mod} \; p$.
\end{theorem}

@@

{\href{http://www.math.Princeton.EDU/~arbooker/nthprime.html}{http://www.math.Pr
inceton.EDU/\~{}arbooker/nthprime.html}} \vskip +7 pt Output of the
nth prime number

@@

\begin{enumerate}
   \item \href{http://www.utm.edu/}{\tt http://www.utm.edu/ }
   \item
\href{http://prothsearch.net/index.html}{\tt
http://prothsearch.net/index.html}
   \item \href{http://www.mersenne.org/prime.htm}{\tt
http://www.mersenne.org/prime.htm }
   \item \href{http://reality.sgi.com/chongo/prime/prime_press.html}{\tt
http://reality.sgi.com/chongo/prime/prime\_press.html }
   \item \href{http://www.eff.org/coop-awards/prime-release1.html}{\tt
http://www.eff.org/coop-awards/prime-release1.html }
\item \href{http://www.eff.org/coop-awards/prime-release1.html}{\tt
%http://www.octocad.demon.co.uk/mersenne/prime.htm }
%   \item \href{http://www.informatik.tu-darmstadt.de/TI/LiDIA/}{\tt
http://www.informatik.tu-darmstadt.de/TI/LiDIA/ }
   \item \href{http://www.math.Princeton.EDU/~arbooker/nthprime.html}{\tt
http://www.math.Princeton.EDU/\~{}arbooker/nthprime.html }
   \item \href{http://www.utm.edu/research/primes/notes/by_year.html}{\tt
http://www.utm.edu/research/primes/notes/by\_year.html }
   \item \href{http://www.cerias.purdue.edu/homes/ssw/cun}{\tt
http://www.cerias.purdue.edu/homes/ssw/cun }
\end{enumerate}

@@

Another name for this type of attack is \index{Man-in-the-middle
attack} ''man-­in-­the-­middle attack''. Users are promised
protection against this type of attack by public­key
certification, which is intended to guarantee the
\index{Authenticity} authenticity of public keys. The most common
certification method is the X.509 standard.

@@

\paragraph{Cyclic groups}
Cyclic groups are those groups $G'$ that possess an element $g$
from which the group operation can be used to generate all other
elements in the group. This means that for each element $a$ in
$G'$ there exists a positive whole number $i$ such that if $g$ is
subject to the operation $i$ times (i.e. ''$g \cdot i$''),
$g+g+...+g = a$ (additive group) or $g^i = g\cdot g \cdots g = a$
(multiplicative group). The element $g$ is the {\em generator} of
the cyclic group --- each element in $G´$ can be generated using
$g$ and the operation.

@@

There are {\em infinite} fields, i.e. the set on which the field is based
contains an infinite number of elements (e.g.: the field of real numbers). And
there are also finite fields, such as ${\mathbb Z}_p = \{0, 1, 2, 3, ..., p-1\}$
, where $p$ is a prime. ${\mathbb Z}_p$ with addition mod $p$ and multiplication
mod $p$ is a finite field.
\index{Field!Characteristic}
\paragraph{Characteristic of a field}
Let $K$ be a field and $1$ be the neutral element of $K$ with
respect to the multiplicative operation ''$\cdot$''. For positive
natural numbers $n$, let us understand $n_1$ to be $n_1 = 1 + 1 +
... + 1$ ($n$ summands and $n_1$ is an element in $K$). If $n_1$
is then unequal to $0$ for all $n>0$, then we call $K$ a field
with characteristic zero. Otherwise, the characteristic of $K$ is
defined to be the smallest positive natural number $p$ for which
$p_1 = 0$ (note: $p$ is then a prime). Comment: The field of real
numbers has the characteristic $0$; the field ${\mathbb Z}_p$ has
the characteristic $p$.

@@

An elliptic curve over ${\mathbb Z}_p$ is defined by an equation of the
following form: $$ y^2 \; ({\rm mod} \; p) = x^3 + ax + b ({\rm mod} \; p) $$
(thus: equality in the field ${\mathbb Z}_p$), where $a, b$ are in ${\mathbb
Z}_p$ and $4a^3 + 27b^2$ mod $p$ is not equal to zero. For fixed chosen numbers
$a$ and $b$ in ${\mathbb Z}_p$, this equation has the pair of solutions $$ {\bf
E} = \left\{(x,y) \left| \begin{array}{c} x {\rm ~and~} y {\rm ~are~in~}
{\mathbb Z}_p {\rm ~and~} \\ y^2  \equiv x^3 + ax + b \; ({\rm mod~} p) {\rm
~and~} \\ 4a^3 + 27b^2 \not\equiv 0 \; ({\rm mod} p)\end{array} \right.
\right\}, $$ i.e. the set ${\bf E}$ consists of all pairs $x$ and $y$ that are a
solution (in ${\mathbb Z}_p$) of the above equation. It must be noted that the
numbers $a, b$ and $p$ determine which pairs $(x,y)$ lie in the set ${\bf E}$.
This means that $a, b$ and $p$ specify this set. The elements $(x,y)$ in ${\bf
E}$ are called points on the elliptic curve. In addition, ${\bf E}$ has one more
element $O$ (the so-called point in infinity). The set ${\bf E}$ is usually
called an elliptic curve. We can now define an operation (also denoted as $+$,
although it is not the standard/usual addition of real numbers) on two elements
in ${\bf E}$ such that the operation delivers an element that also lies in ${\bf
E}$. The set ${\bf E}$ is therefore closed under the operation $+$. We can show
that ${\bf E}$ is a group. The neutral element of the group ${\bf E}$ is the
point in infinity $O$. Thus, for every two points $(x_1,y_1)$ and $(x_2,y_2)$ on
the elliptic curve ${\bf E}$, there exists a point $(x_3,y_3)$ on ${\bf E}$ such
that the operation $+$ complies with the following: $(x_1,y_1) + (x_2,y_2) =
(x_3,y_3)$. Under certain circumstances, these points may also be equal to the
point in infinity. Thus, when we speak of a point $P$ on an elliptic curve ${\bf
E}$, we mean that $P = (x,y)$ and $(x,y)$ lies in the set ${\bf E}$. Any two
points on an elliptic curve specified by $a, b$ and $p$ can therefore be added
and the result is a point that also lies on the same elliptic curve. We must
note that ${\bf E}$ can have the following meanings:
\begin{itemize}
   \item the set ${\bf E}$ of solutions pairs $(x,y)$ for an equation including the point $O$
   \item the group ${\bf E}$ (with the operation ''addition of $(x_1,y_1)$ and $(x_2,y_2)$'')
   \item the elliptic curve ${\bf E}$ (which is actually the same as the group ${\bf E}$)
 \end{itemize}

@@

For cryptography, the important fact is that, for very large
numbers, it appears to be extremely difficult to use a given point
$Q$ on an elliptic curve to determine which two points have to be
added to obtain $Q$. For large numbers $a, b$ and $p$ ($p$, for
example, has a length of more than $160$ bits), the computer can
easily add the point $P$  $m$ times after another, i.e. to
determine the point $P + P + \cdots + P = Q$ in an incredibly
short space of time (in a few fractions of a second) ($m$ summands
$P$). Rather than $P + P + \cdots + P = Q$ ($m$ summands $P$) we
also write $mP = Q$. If we have a point $P$ and a point $Q$, which
both lie on the same elliptic curve, no procedure is known that
enables us --- within an acceptable space of time --- to determine
the number $m$ (assuming it actually exists) for which $mP = Q$.
This is referred to as the ''elliptic curve discrete logarithm
problem'' (abbreviated to ECDLP).

We must note that not all elliptic curves are equally secure. This
means that we must choose the parameters $a$ and $b$ carefully
when defining a curve. For certain classes of elliptic curves, it
is possible to solve the ECDLP easier than in the general case.
Cryptographically unsuitable elliptic curves are called abnormal
curves (these are curves over ${\mathbb Z}_p$ for which the set
${\bf E}$ has precisely $p$ elements) and the supersingular curves
(the curves for which we can reduce the calculation of the ECDLP
to calculating the ''standard'' discrete logarithm in other finite
fields, i.e. simplify the calculation). There are therefore
cryptographically good and bad curves. However, for given
parameters $a$ and $b$ we can, with some difficulty, establish
whether or not the resulting elliptic curve is useful for
cryptographic purposes. The curves used in cryptography are
usually provided by experts. They ensure that the elliptic curves
they classify as secure satisfy the current security requirements.
For secure curves, the parameter $p$ determines how long it takes
to solve the ECLDP on this curve. The larger the parameter $p$,
the longer it takes to solve the problem. Experts recommend a bit
length of over $200$ bits for the parameter $p$. This makes it
clear why elliptic curves are so interesting for cryptography.
Because the parameter $p$ also determines the time required to
perform the signature/encryption procedure when using elliptic
curves in cryptography. The time taken to generate a pair of keys
also depends on $p$. Thus, small values (few bits) are desirable
here (in order to minimise the run times for the procedures);
however, the required security must still be maintained. For
example, with a length of $200$ bits for $p$, a {\em good}
elliptic curve is just as secure as an \index{RSA!module} RSA
module of over $1024$ bits in length (at least according to the
current state of research). The reason for this is that the
quickest algorithms for solving the {\em elliptic curve discrete
logarithm} problem have an exponential run time --- unlike the
subexponential run times that the best factorisation algorithms
currently have (number sieve, quadratic sieve or factorisation
with elliptic curves). Therefore, the parameters for cryptographic
procedures based on the problem of {\em factorising whole numbers}
must be greater than the parameters for cryptographic procedures
based on the ECDL problem.

@@

The {\em elliptic curve discrete logarithm problem} (ECLDP) forms the basis for
elliptic-curve cryptography. Various signature procedures are based on this.
What they have in common is how they use the public parameters and how these
parameters are used to generate secret and public keys:
\begin{itemize}
    \item The parameters of the elliptic curve E, i.e. a prime number $p$ that
determines over which field ${\mathbb Z}_p$ the elliptic curve ${\bf E}$ is
defined, as well as the two numbers $a$ and $b$ in ${\mathbb Z}_p$.
    \item A point $G=(x,y)$ that lies on the elliptic curve ${\bf E}$.
    \item A prime number $r < p$ for which $rG=O$ (i.e. $G$ added $r$ times
gives the neutral element of the group ${\bf E}$) and $r$ is a factor of $\#{\bf
E}$ (where $\# {\bf E}$ is the number of elements in the set ${\bf E}$). The
point $G$ therefore has the order $r$ and is the generator of a cyclic subgroup
of ${\bf E}$ with the order $r$.
    \item The number $k = \#{\bf E}/r$ ($k$ is called the {\em cofactor}).
\end{itemize}
The parameters $a, b, p, G, r$ and $k$ listed above are called
domain parameters. They determine on which elliptic curve ${\bf
E}$ and in which cyclic subgroup of ${\bf E}$ a signature
procedure has been ''used''.
